Cost-Benefit Analysis

### 2023-05-31 workflow/notebooks/cost-benefit-analysis.Rmd

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

- use values from confusion matrix for one representative model on a test set. or get average precision? choose a particular sensitivity level and see the range in precision?
  - Number needed to screen (NNS) - the number of alerted patients the models must flag to identify 1 true positive. 
  - Number needed to treat (NNT) - the number of true positive patients one must treat for 1 patient to benefit from the treatment.
  - Number needed to benefit (NNB = NNS x NNT) - how many patients must be screened for 1 patient to benefit.
- assumptions
  - cost of a non-severe case
  - cost of a severe case
    - days in icu, colectomy
  - cost of possible treatments:
    - abx: metronizadole, vancomycin, fidaxomycin
    - fmt, bezlotoxumab (monoclonal antibodies)

@@

@@

```{r decision-curve}
library(here)
library(tidyverse)
thresh_dat <- read_csv(here('results','thresholds_results_aggregated.csv'))
thresh_dat %>% 
  filter(dataset == 'int',outcome=='attrib',seed==100) %>% 
  ggplot(aes(decision_threshold, net_benefit, 
             color = strategy, linetype = strategy)) +
  geom_line(alpha = 0.6) +
  theme_sovacool()
```

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      fig.path = 'figures/')
```
```{r load}
library(here)
library(schtools)
library(tidyverse)
thresh_dat <- read_csv(here('results','thresholds_results_aggregated.csv')) %>% 
  mutate(predicted_pos_frac = (tp + fp) / (tp + fp + tn + fn))
treat_all <- thresh_dat %>% 
  filter(strategy == 'all') %>% 
  select(strategy, decision_threshold, net_benefit, dataset, outcome) %>% 
  unique() %>% 
  pivot_wider(names_from = strategy, values_from = net_benefit, 
              names_prefix = "nb_treat_")
treat_none <- thresh_dat %>% filter(strategy == 'none') %>% 
  select(strategy, decision_threshold, net_benefit, dataset, outcome) %>% 
  unique() %>% 
  pivot_wider(names_from = strategy, values_from = net_benefit, 
              names_prefix = "nb_treat_")
thresh_dat <- thresh_dat %>% 
  filter(strategy == 'model') %>% 
  left_join(treat_all, by = join_by(decision_threshold, dataset, outcome)) %>% 
  left_join(treat_none, by = join_by(decision_threshold, dataset, outcome))
```

- use values from confusion matrix for one representative model on a test set. or get average precision? choose a particular sensitivity level and see the range in precision?
- assumptions
  - cost of a non-severe case
  - cost of a severe case
    - days in icu, colectomy
  - cost of possible treatments:
    - abx: metronizadole, vancomycin, fidaxomycin
    - fmt, bezlotoxumab (monoclonal antibodies)

@@

## percentile of risk

95th percentile of risk: risk threshold where action is taken in 5% of patients.

```{r risk-pct}
roc_risk_pct <- thresh_dat %>% 
  mutate(Specificity = round(Specificity, 2)) %>%
  group_by(Specificity, dataset, outcome) %>%
  summarise(
    mean_sensitivity = mean(Sensitivity),
    mean_pred_pos_frac = mean(predicted_pos_frac)
   ) %>% 
  ungroup() %>%
  group_by(dataset, outcome) %>%
  mutate(diff_95th_pct = abs(mean_pred_pos_frac - 0.05)) %>%
  slice_min(diff_95th_pct)
prc_risk_pct <- thresh_dat %>% 
  mutate(Recall = round(Recall, 2)) %>%
  group_by(Recall, dataset, outcome) %>%
  summarise(
    mean_precision = mean(Precision),
    mean_pred_pos_frac = mean(predicted_pos_frac)
   ) %>% 
  ungroup() %>%
  group_by(dataset, outcome) %>% 
  mutate(diff_95th_pct = abs(mean_pred_pos_frac - 0.05)) %>%
  slice_min(diff_95th_pct)

# confusion matrix values for 95th percentile of risk
confmat_95th_roc <- thresh_dat %>% 
  mutate(Specificity = round(Specificity, 2)) %>%
  inner_join(roc_risk_pct, by = join_by(Specificity, dataset, outcome)) %>% 
  mutate(diff_mean_sens = abs(mean_sensitivity - Sensitivity)) %>% 
  group_by(dataset, outcome) %>% 
  slice_min(diff_mean_sens) %>% 
  select(-seed) %>% 
  unique() %>% 
  slice_max(net_benefit) %>% 
  select(outcome, dataset, Precision, Recall, Specificity, tp,fp,tn,fn,
         net_benefit, nns, decision_threshold)

confmat_95th_prc <- thresh_dat %>% 
  mutate(Recall = round(Recall, 2)) %>% 
  inner_join(prc_risk_pct, by = join_by(Recall, dataset, outcome)) %>% 
  mutate(diff_mean_prec = abs(mean_precision - Precision)) %>% 
  group_by(dataset, outcome) %>% 
  slice_min(diff_mean_prec) %>% 
  select(-seed) %>% 
  unique() %>% 
  slice_max(net_benefit) %>% 
  select(outcome, dataset, Precision, Recall, Specificity, tp,fp,tn,fn,
         net_benefit, nns, decision_threshold)
confmat_95th_prc
```

## decision curve based on cost?

@@

```{r decision-curve}
max_thresholds <- thresh_dat %>% 
  filter(strategy == 'model') %>% 
  group_by(dataset, outcome, strategy, decision_threshold)  %>% 
  summarize(mean_nb = mean(net_benefit)) %>% 
  filter(mean_nb <= 0) %>% 
  slice_min(mean_nb) %>% 
  ungroup() %>% 
  mutate(max_thresh = decision_threshold) %>% 
  select(dataset,outcome,max_thresh)

thresh_dat %>% 
  left_join(max_thresholds, by = join_by(dataset, outcome)) %>% 
  filter(decision_threshold <= max_thresh) %>% 
  group_by(dataset, outcome, decision_threshold) %>% 
  summarize(mean_nb = mean(net_benefit),
            nb_treat_all = mean(nb_treat_all),
            nb_treat_none = mean(nb_treat_none)) %>% 
  ggplot(aes(x = decision_threshold)) +
  geom_line(aes(y = mean_nb, color = outcome), alpha = 0.6) +
  geom_line(aes(y = nb_treat_all), linetype = 'dotted') +
  geom_line(aes(y = nb_treat_none), linetype = 'dashed') +
  scale_y_continuous(limits = c(-0.1,0.1)) +
  facet_grid(dataset ~ outcome, scales = 'free') +
  theme_sovacool()
```
